---
title: "Hierarchical Clustering"
description: |
  For an assignment in ESM 244 I explored the multivariate clustering of water chemistry among creeks in Santa Barbara Country. I used hierarchical clustering by complete linkage to create a dendrogram showing my findings.
author:
  - name: Laurel Wee
    url: {}
date: 03-15-2021
output:
  distill::distill_article:
    self_contained: false
    code_folding: expand 
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(tidyverse)
library(here)
library(naniar)
library(NbClust)
library(ggdendro)
library(dendextend)
```


```{r}
# Read in steam chemistry data from https://drive.google.com/file/d/16rYLBi-CgvjcWhOsX1SLdD9HHUMP9m2l/view?usp=sharing
stream_chem <- read.csv(here("data","sbc_lter_registered_stream_chemistry - sbc_lter_registered_stream_chemistry.csv")) %>% 
  na_if(-999.0)%>% # turn `-999.0` values into NA
  select(site_code, ends_with("uM"))
         #nh4_uM,no3_uM,po4_uM) # Select `site_code` mirco-molar concentration columns
```

```{r}
# Make a new dataframe that includes the average of each measurement by site
stream_avg <-stream_chem %>% 
  group_by(site_code) %>% 
  summarise_all(mean, na.rm=TRUE) %>% 
  mutate(site_name= # Add column with full site names from meta data https://portal.edirepository.org/nis/metadataviewer?packageid=knb-lter-sbc.6.16
         case_when(
           site_code == "AB00" ~ "Arroyo Burro",
           site_code == "AT07" ~ "Atascadero Creek",
           site_code == "BC02" ~ "Bell Canyon Creek",
           site_code == "DV01" ~ "Devereaux Creek",
           site_code == "GV01" ~ "Gaviota",
           site_code == "HO00" ~ "Arroyo Hondo",
           site_code == "MC00" ~ "Mission Creek (Montecito)",
           site_code == "MC06" ~ "Mission Creek (Rocky Nook)",
           site_code == "ON02" ~ "San Onofre Creek",
           site_code == "RG01" ~ "Refugio Creek",
           site_code == "RS02" ~ "Rattlesnake Creek",
           site_code == "SP02" ~ "San Pedro Creek",
           site_code == "TO02" ~ "Tecolote Creek"))



# Scale the averaged dataframe
stream_scale <- stream_avg %>% 
  select(ends_with("uM")) %>% 
  scale()
# Make `site_code` the rownames of the averaged data frame
rownames(stream_scale)<-stream_avg$site_name 

# Make a subset of data with only the sites that have all the uM data recorded
stream_full<- stream_scale %>% 
  na.omit()

```

```{r}
# Compute dissimilarity values (Euclidean distances):
euc_distance <- dist(stream_full, method = "euclidean")

#View(euc_distance)

# Hierarchical clustering (complete linkage)
hc_complete <- hclust(euc_distance, method = "complete" ) 

# Make a dendrogram visual
ggdendrogram(hc_complete,
             rotate = TRUE)+
  theme_minimal()+
  theme(panel.background=element_rect("lightblue1"))+
  labs(x="Creek Site", y="Height", title="Water Chemistry of Santa Barbara County Creeks (WEE)")
  
```

### Summary
Only nine out of 13 creek sites in Santa Barbara County had micro molar concentration data for water chemistry across eight variables. Mission Creek at Montecito and Arroyo Burro had the least dissimilarities among all the sites while San Onofre Creek and Gaviota has the greatest dissimilarities. 

####  Data Citation:
Santa Barbara Coastal LTER and J. Melack. 2019. SBC LTER: Land: Stream chemistry in the Santa Barbara Coastal drainage area, ongoing since 2000 ver 16. Environmental Data Initiative. https://doi.org/10.6073/pasta/67a558a24ceed9a0a5bf5e46ab841174. 

